Bayesian Evidence Synthesis:opioid crisis

### 2020-03-11 Rcodes/JAGS_overdoses.Rmd
N uncertain work -ing

```{r}
visualization(my_list_unbiased, x_a=T)
visualization(my_list_under,x_a= T)
visualization(my_list_over,x_a=T)
```
############# Below here is for the purpose of debugging ##################

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
## Trim here and make a function and give us the plots.
df <- read.csv('./under_p_a_data.csv')
df$X <- NULL 
df$bias <- NULL
# obtain the (biased) data 
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

@@

            )
# run the model 
head(df)
```

```{r}
chains=2
under.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)

```{r}
# inits = list()

# get the samples of O_t, p_a
params= c('o_t','p_a')
samples <- coda.samples(under.model, params, n.iter = 1000)
pst_mtx = as.matrix(samples)
head(pst_mtx)
# tidy p_a
p_a <- pst_mtx[,13]
head(p_a)
```

## tidy o_t
pst_mtx = as.matrix(samples)
temp = pst_mtx[,1:12]
colnames(temp) <- seq(1,12)
df_o_t <- as.data.frame(temp)
trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
# real values of the data set from pymc3 samples
o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
# factorizing the months for box plot visualization
o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
# boxplots of o_t
ggplot()+
  # boxplot from the trace 
  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
  # real values as red dots
  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')
```

# tidy u_t pp samples
df_u_t <- data.frame()
for (i in 1:m) {
  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
  df_u_t=rbind(df_u_t,obs)
}
colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
# boxplots of u_t
ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
```
```{r}
## tidy x_a
m=length(p_a)
df_x_a <- vector() 
for (i in 1:n_T) {
  obs <- rbinom(m,n_a,p_a)
  df_x_a <- cbind(df_x_a,obs)
}

colnames(df_x_a) <- seq(1:12)
df_x_a <- as.data.frame(df_x_a)
ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))

>>>>

```{r}
visualization(my_list_unbiased, x_a=T)
visualization(my_list_under,x_a= T)
visualization(my_list_over,x_a=T)
```
# Uncertainty of N
Now we suppose the population size of BC, N, is unknown. We may have some information about the population size and transform it into a prior distribution. Let's assume from the point that we don't have a good understanding of the sample size. But N should be at least bigger than the sample size of survey, the ambulance attended overdoses. We have 12 data points of $U_t$. We also have 12 data points of $x_A$.

$$N \geq max(\vec{U_t}, \vec{x_A})$$

Where $\vec{U_t}, \vec{x_A}$ are the vectors of $U_t$ and $x_A$.

Then, we need a theoretical maximum value of N. BC is a province of Canada. For now, since I do not have a good idea about the maximum possible number of N, let's just say two times of the real value is the maximum.

$$N \in  \{a, a+1, a+2, ..., b\} \\ \textbf{where } a= \max(\vec{U_t}, n_A), b= 2 \times \textbf{real value of  N}$$

However, in the model file below, we will use a prior such that 
$$N \in  \{0,1,2, ..., b\} \\ \textbf{where }  b= 2 \times \textbf{real value of  N}$$ since we don't know how to construct such prior as described below.

```{r Enter the models}
cat("model{
## define the priors

# Uncertain N takes in to the model
N ~ dcat(p_i)

}", file='N_uncertain_model.txt')
```

Pre-set variables.  
```{r Defining Values}
p_i <- seq(1:(2*N))
n_T <- length(df$o_t)
n_a <- 1000
# N <- 10000
u_t <- df$u_t
x_a <- df$x_a
```

Define the list providing the values of the variables and the parameters for the priors of the model.   
```{r from variable name to \'variable_name\'}
dat.N <- list(# prior for N
            'p_i' = p_i,
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

@@

            # likelihood 
            'u_t'=u_t, # giving data 
            'x_a'=x_a,   # giving data
            'n'= n_T, # total months 12
            'n_a'=n_a # survey size 1000

            )
```

``` {r run the first model}
chains=2
# inits = list()
simple.N.model <- jags.model(file='N_uncertain_model.txt',
                          data=dat.N,
                          n.chains = chains)
simple.N.model
```

```{r}
test <- coda.samples(model = simple.N.model,variable.names = c('o_t','p_a','N'), n.iter = 1000)
interations = 1000
burnin= floor(interations/2)
summary(window(test), start=burnin)
head(test)
```

```{r get the samples of N and show a distribution}
pst_mtx.N = as.matrix(test)
head(pst_mtx.N)
N <- pst_mtx.N[,1]
```
```{r}
ggplot(data = as.data.frame(N), aes(x= N)) +geom_histogram()+ geom_vline(xintercept =10000)
```

It took a lot of time to fit the bayesian model this time. We have overestimated N value whose mode is around 17000 while the real value is 10000. Unif(0,20000) has 10000 as its mean. I guess the b was too large. 
I will wait for the reply from Paul. 
```{r}
plot(test)
```

### 2020-03-11 latex/newpaper.tex
N uncertain work -ing

\normalsize 
The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$. The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality. The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.  \\

@@

\normalsize 
There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets are simulated as follows. 
The true value of $p_A$ was set $p_A=0.8$ for the survey data.
It is assumed that the data was collected for a year (t=1,2,3, ..., 12) and 
$x_t$ values were independentally generated from the Binomial distribution (\ref{ambulance}).  In terms of overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
The vector of $O_t$ was generated following the overdose model (\ref{overdose}). The vector of $U_t$ was generated from the Binomial relation of the two variables ($\ref{over_amb}$). The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12). [Delete: Then the 12 sueveys were combined into a single survey since $p_A$ is assumed fixed across time in this contexts.] Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\

@@

\begin{equation}
\label{nonin_prior_amb}
p(p_A) \sim Beta(1,1)
\text{			noninformative prior of ambulance model}
\end{equation}

\begin{equation}
\label{noninprior_over}
\left.\begin{aligned}
\mu \sim U(-10,0)\\
\sigma \sim U(0,5)
\end{aligned}\right\} 
\text{			noninformative prior of overdose model}
\end{equation}

@@

>>>>

\normalsize 
The first simulation simplifies the assumptions of variables as much as possible; we assumed $N= 10000, n_{A}=1000$.  \\

@@

\normalsize 
There exist two data sets; survey data ($n_A, x_A$), and ambulance attended overdose data ($U_t$). The two data sets are simulated as follows. 
The true value of $p_A$ was set as $p_A=0.8$ for the survey data. [n_A? the setup described here above has a single x ]
It is assumed that the data was collected for a year (t=1,2,3, ..., 12) and 
$x_t$ values were independentally generated from the Binomial distribution (\ref{ambulance}).  In terms of overdose data, It is assumed that the true values of parameters for overdose model were $\mu=\log0.05,\ \sigma=1$.
The vector of $O_t$ was generated following the overdose model (\ref{overdose}). The vector of $U_t$ was generated from the Binomial relation of the two variables ($\ref{over_amb}$). The two generated vectors have the same length with the survey data (t=1,2,3, ..., 12).  Note that only $U_t$ and $x_t$ are known as the likelihood and $p_A$ needs to be estimated first so as to estimate $O_t$ which is the ultimate interest of the research.\\

@@

\begin{equation}
\label{nonin_prior_amb}
p(p_A) \sim Beta(1,1)
\text{			noninformative prior for ambulance model}
\end{equation}

\begin{equation}
\label{noninprior_over}
\left.\begin{aligned}
\mu \sim U(-10,0)\\
\sigma \sim U(0,5)
\end{aligned}\right\} 
\text{			noninformative prior for overdose model}
\end{equation}

@@

\subsection{Variying the Assumptions}
The assumptions will change later to see the impact of the likelihood over the posterior distributions of variables of interest; the total population size for a region, N, could vary over time or it can be staritified for a better realization of the reality. The survey size could also vary such as $n_{A}=100$ or $n_{A}=10000$ in later examples.
